---
title: "classification"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions before/during reading:

1. maths behind logistic function? 

2. what is log-odds/logit? 
  - "logistic regression model, increasing X by one unit changes the log odds by $\beta_1$ or equivalently multiplies the odds by $e^{\beta_1}$"
  - "a one-uni increase in balance is associated withan increase in the log odds of default by 0.0055 units" (p.134)... what does that mean?

3. "The estimated intercept in Table 4.1 is typically not of interest" (p.134)...why?

4. Keen to talk through this example: "a student is riskier than a non-student if no information about the student's credit card balance is available. However, that student is less risky than a non-student *with the same credit card balance*"

5. maximum likelihood vs least squares?

# Breast Cancer classification problem
```{r}
set.seed(22)
library(mlbench)
library(tidyverse)
library(janitor)
library(caTools)

data(BreastCancer)

bc_data_raw <- BreastCancer %>%
  as_tibble() %>%
  clean_names()

str(bc_data_raw)

sum(is.na(bc_data_raw))

# going to reduce data set down to 4 variables of interest
bc_data <- bc_data_raw %>%
  na.omit() %>%
  select(cl_thickness, cell_size, cell_shape, marg_adhesion, class)

str(bc_data)

# let's reduce the number of factor levels from 10 to 3
table(bc_data$cl_thickness) 
table(bc_data$cell_size) 
table(bc_data$cell_shape) 
table(bc_data$marg_adhesion) 

bc_data_clean <- bc_data %>%
  mutate(across(cl_thickness:marg_adhesion, ~ case_when(.x %in% c(1,2,3) ~ "Low", 
                                   .x %in% c(4,5,6,7) ~ "Medium",
                                   .x %in% c(8,9,10) ~ "High"))) %>%
  na.omit()

str(bc_data_clean)

bc_data_clean$cl_thickness <- factor(bc_data_clean$cl_thickness, levels = c("Low", "Medium", "High"))
bc_data_clean$cell_size <- factor(bc_data_clean$cell_size, levels = c("Low", "Medium", "High"))
bc_data_clean$cell_shape <- factor(bc_data_clean$cell_shape, levels = c("Low", "Medium", "High"))
bc_data_clean$marg_adhesion <- factor(bc_data_clean$marg_adhesion, levels = c("Low", "Medium", "High"))
```

# Data visualisation 
```{r}
ggplot(bc_data_clean, aes(cl_thickness)) +
  geom_bar(aes(fill=class), color="black") +
  theme_bw()

ggplot(bc_data_clean, aes(cell_size)) +
  geom_bar(aes(fill=class), color="black") +
  theme_bw()

ggplot(bc_data_clean, aes(cell_shape)) +
  geom_bar(aes(fill=class), color="black") +
  theme_bw()

ggplot(bc_data_clean, aes(marg_adhesion)) +
  geom_bar(aes(fill=class), color="black") +
  theme_bw()


```


# Split into training and testing data
```{r}
bc_data_clean$split <- sample.split(bc_data_clean$class, SplitRatio = 0.7)

bc_train <- bc_data_clean %>%
  filter(split == TRUE)

bc_test <- bc_data_clean %>%
  filter(split == FALSE)

bc_model <- glm(class ~ ., family = binomial(link = "logit"), data = bc_train)
summary(bc_model)

bc_test$predicted_class <- predict(bc_model, newdata = bc_test, type = "response") # not sure what this warning is?? 

# confusion matrix
confusion_mat <- table(bc_test$class, bc_test$predicted_class > 0.5) # compares actual class to predicted class 
confusion_mat

acc <- (confusion_mat[1,1]+confusion_mat[2,2])/sum(confusion_mat)
# model is 96% accurate
```



