---
title: "regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions I had before/during reading:
- what is the difference between correlation and linear regression? when should you use one vs the other?
- is mean centering the same as standardizing coefficients? when should you do either?
- what are people's opinions of R^2 vs adjusted R^2?    
  - R^2 is more interpretable whereas adjusted accounts for bias 
- are some assumptions more important than others? do people actually check the assumptions?
- what does is mean to say "residuals that are independent of each other"?
- what are the different types of residuals? regular vs standadised vs studentised?
- what is an ideal workflow for checking your model and all the assumptions?
- what to do when specific assumptions are violated?

- [datasets in R](https://machinelearningmastery.com/machine-learning-datasets-in-r/)
- we'll use: Longley’s Economic Regression Data
```{r}
library(tidyverse)
library(janitor)
library(car)
library(mlbench)

theme_set(theme_bw())
```

# simple linear regression
- Boston Housing dataset 
- first, let's look at relationships between different variables
```{r}
data(BostonHousing)
?BostonHousing

boston <- BostonHousing %>%
  as_tibble() 

pairs(boston, pch=19, lower.panel = NULL)
```

- pretty clear relationship between rm (avg number of rooms per dwelling) and medv (median value of owner-occupied homes)

- dis (distance to 5 employment centers) and nox (nitric oxides concentration) perhaps quadratic relationship??
  - this seems more interesting so let's look at this 
  - also good exercise because a lot of times our data won't be linear
```{r}
ggplot(boston, aes(dis, nox)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) # not surprising, but not a very good fit 

m1 <- lm(dis ~ nox, data=boston)
m1
summary(m1)

names(m1)
m1$coefficients # or coef(m1)

predict(m1, data.frame(nox=c(.3,.5)),
        interval="confidence") 

```

# check assumptions: what to look for in each?
- plot 1: residuals and fitted values (want to see a straight line?)
- plot 2: normal QQ
- plot 3: scale location
- plot 4: residuals vs leverage
```{r}
plot(m1)

which.max(hatvalues(m1)) # which observation has the largest leverage
# what about 354? is that just an outlier?
```

# multiple linear regression
```{r}
m2 <- lm(dis ~ ., data=boston) # all predictors
summary(m2)

summary(m2)$r.sq

vif(m2) # help determine if predictors are too highly correlated
# some numbers are large what to do next? try to remove collinearity?

m3 <- lm(dis ~ . - tax - lstat, data=boston) # not sure which variables to remove? systematic way to do this?
vif(m3) 
```

# non linear transformation
- when transforming predictors so it's ^2, you need to wrap it in `I()` because ^ has a special meaning in a formula 
- the fact that the transformed predictor (quadratic term) is significant, near 0, suggests it leads to an improved model
- can do anova() to quantify the extent to which the quadratic fit is superior to the linear fit
- can use the poly() to test multiple polynomial terms at once
```{r}
m4 = lm(dis ~ nox + nox^2, data=boston)
summary(m4)

m5 = lm(dis ~ nox + I(nox^2), data=boston)
summary(m5)

anova(m1, m5)

plot(m5)

# what about ^3? or ^4
m6 = lm(dis ~ poly(nox,5), data=boston)
summary(m6)

```


