---
title: "regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions I had before/during reading:
- what is the difference between correlation and linear regression? when should you use one vs the other?
- is mean centering the same as standardizing coefficients? when should you do either?
- what are people's opinions of R^2 vs adjusted R^2?    
  - R^2 is more interpretable whereas adjusted accounts for bias 
- are some assumptions more important than others? do people actually check the assumptions?
- what does is mean to say "residuals that are independent of each other"?
- what are the different types of residuals? regular vs standadised vs studentised?
- what is an ideal workflow for checking your model and all the assumptions?
- what to do when specific assumptions are violated?

- [datasets in R](https://machinelearningmastery.com/machine-learning-datasets-in-r/)
- we'll use: Longley’s Economic Regression Data
```{r}
library(tidyverse)
library(janitor)
library(car)

theme_set(theme_bw())
```

# simple linear regression
```{r}
data(longley) 

?longley

longley <- longley %>%
  as_tibble() %>%
  clean_names()
  
ggplot(longley, aes(gnp, employed)) +
  geom_point() +
  geom_smooth(method = "lm", se=FALSE) +
  xlim(0, 700) +
  ylim(50,75)

m1 <- lm(gnp ~ employed, data=longley)
m1
summary(m1)

names(m1)
m1$coefficients # or coef(m1)

predict(m1, data.frame(employed=c(50,55,80)),
        interval="confidence")

plot(m1)

which.max(hatvalues(m1)) # which observation has the largest leverage 
```

# multiple linear regression
```{r}
m2 <- lm(gnp ~ employed + year, data=longley)
summary(m2)

m3 <- lm(gnp ~ ., data=longley) # all predictors
summary(m3)

summary(m3)$r.sq

vif(m3) # why are numbers so high? collinearity? 

m4 <- lm(gnp ~ . - gnp_deflator - unemployed, data=longley) # all predictors
vif(m4) # still high
```


